Analysis of competitive surfing tournaments with generalized Bradley-Terry likelihoods

### 2022-03-21 inst/surfing_jsa_rebuttal.tex

\textcolor{blue}{See the table of results at the end of this document.
  We see that the wildcards participate in very few venues: of the 25
  wildcards, 20 appear at only a single venue, four at two venues, and
  only one (Willcox) at three venues.  Further, the wildcards perform
  poorly: of the 31 ranked scores, only 5 are better than last or
  second to last.  Maximum likelihood Bradley-Terry strengths for such
  competitors are either exactly zero or very small (here $\sim
  10^{-5}$) and are neither interesting from a surfing lore
  perspective, nor susceptible to meaningful statistical
  inference.\newline Quite apart from the fact that our focus was on
  the top-performing competitors, including such datapoints is
  problematic for several reasons.  Firstly, numerical optimization
  techniques tend to struggle as one increases the number of players:
  one is necessarily working with a higher-dimensional problem,
  leading to poorly defined maxima [the Jacobian becomes very small];
  in addition, local maxima become more numerous and difficult to
  identify.\newline We have added a brief summary of this to the
  manuscript.}

\subsection*{Direct comparison on particular waves}

One other comment is based around the statements around judging in the
second last paragraph before section 1.1. Judging is based on the
heats conditions, as surfing is performed within an ever changing
environment. Therefore, to try and compare between surfers throughout
a competition is really not applicable. As each heat, and each wave is
different from the previous. So a direct comparison throughout a
competition is somewhat unfair for the competitors. As for the claim
around abilities, is this not subjective, as each performer is judged
upon their surfing prowess on that particular wave in that particular
heat.

\textcolor{blue}{Competition is indeed performed within an ever
  changing environment.  Each wave is indeed different from the
  previous.  We believe that these facts are the reason for the
  scoring system adopted in WSL tournaments.  We point out that the
  environment during any individual heat is common to all competitors
  in the water, so comparison between the competitors is fair, in the
  sense that no competitor can justifiably claim that he was placed at
  a disadvantage by wave quality.  All the competitors in the water
  have the same opportunity to catch the same waves (subject to wave
  priority rules, but these rules are fair in the same sense).  We
  would contend that a competitor's ability to choose an opportune
  wave is part of the skill of surfing, and is certainly necessary for
  tournament success.}

\textcolor{blue}{We use the word ``subjective" to mean ``personal
  aesthetic valuation"; we wished to emphasise that pursuits such as
  surfing are scored on judgements, rather than objective physical
  measures such as timings or heights as seen in, for example, 100m
  sprinting or archery.}

### 2022-03-21 inst/surfing_jsa_rebuttal.tex

\documentclass[12pt]{article}
\usepackage{xcolor}
\begin{document}

\section*{MS596 ``Analysis of competitive surfing tournaments with generalized Bradley-Terry likelihoods" by Driver and Hankin: rebuttal}

Below, the reviewer's comments are in black, and our replies to the
issues are in \textcolor{blue}{blue}.  We have indicated changes to
the manuscript where appropriate.

Short story: we have accommodated all the comments with rewording and
clarification.  The resulting document is, we believe, stronger and
more scholarly than before and we recommend it to you.

@@

\textcolor{blue}{We have clarified that we are studying the men's
  championship}

\subsection*{Ignoring wildcards}

Why are you ignoring the "wildcards". No real justification for
this. Especially as some of the wildcards competed in more than one of
the competitions. So some form of statement to validate why you omit
the wildcards would be good.

\textcolor{blue}{See the table of results at the end of this document.
  We see that in general, a wildcard competitor participates in very
  few venues: of the 25 wildcards, 20 appear at only a single venue,
  four at two venues, and only one (Willcox) at three venues.
  Further, the wildcards perform poorly: of the 31 ranked scores, only
  5 are better than last or second to last.  Maximum likelihood
  Bradley-Terry strengths for such competitors are either exactly zero
  or very small (here $\sim 10^{-5}$) and are neither interesting from
  a surfing lore perspective, nor susceptible to meaningful
  statistical inference.\newline Quite apart from the fact that our
  focus was on the top-performing competitors, including such
  datapoints is problematic for several reasons.  Firstly, numerical
  optimization techniques tend to struggle as one increases the number
  of players: one is necessarily working with a higher-dimensional
  problem, leading to poorly defined maxima [the Jacobian becomes very
    small]; in addition, local maxima become more numerous and
  difficult to identify.\newline We have added a brief summary of this
  to the manuscript.}

@@

\subsection*{Direct comparison on particular waves}

One other comment is based around the statements around judging in the
second last paragraph before section 1.1. Judging is based on the
heats conditions, as surfing is performed within an ever changing
environment. Therefore, to try and compare between surfers throughout
a competition is really not applicable. As each heat, and each wave is
different from the previous. So a direct comparison throughout a
competition is somewhat unfair for the competitors. As for the claim
around abilities, is this not subjective, as each performer is judged
upon their surfing prowess on that particular wave in that particular
heat.

\textcolor{blue}{Competition is indeed performed within an ever
  changing environment.  Each wave is indeed different from the
  previous.  We believe that these facts are the reason for the
  scoring system adopted in WSL tournaments.  We point out that the
  environment during any individual heat is common to all competitors
  in the water, so comparison between the competitors is fair, in the
  sense that no competitor can justifiably claim that he was placed at
  a disadvantage by wave quality.  All the competitors in the water
  have the same opportunity to catch the same waves (subject to wave
  priority rules, but these rules are fair in the same sense).  We
  would contend that a competitor's ability to choose an opportune
  wave is part of the skill of surfing, and is certainly necessary for
  tournament success.}

\textcolor{blue}{We use the word ``subjective" to mean ``personal
  aesthetic valuation"; we wished to emphasise that pursuits such as
  surfing are scored on judgements, rather than objective physical
  measures such as timings or distances as seen in, for example, 100m
  sprinting or archery.}

@@

\subsection*{Brazilian preference}

Your hypothesis on Brazilians preference for reef breaking waves is
based on what?

\textcolor{blue}{We have clarified the issue and corrected an
  unfortunate typographical error.  The new manuscript reads: ``one
  suggestion (Burgess 2020, Ho 2021) is that Brazilian surfers tend to
  have skillsets that favour point and beach break wave types, readily
  accessible to these surfers' native Brazil; see (Scarfe 2003) for a
  scientific overview of surf break characteristics, and (Butt 2004)
  for a surfer's perspective''.  In short, we suggest that Brazilians
  {\em do not} like reef breaking waves.}

@@

\subsection*{Clarification}

Other than that, an interesting paper. Just feel you need to clarify a
few of the points above to provide readers a clearer viewpoint of the
sport and understanding to the importance of your study. As I found it
quite interesting. Just some claims and anomalies related to
competitive surfing.

\section*{Detailed rebuttal: Reviewer \#2}

{\bf manuscript:} The World Surf League (WSL) is the main governing
  body of surfing competitions (World Surf League 2021).

{\bf comment:} Not exactly. What about ISA (world surfing games, Olympics, national
championships, regional championships).

\textcolor{blue}{{\bf response:} We have clarified that the WSL is the body for professional surfers}

\rule{0mm}{10mm}

{\bf manuscript:} The WSL conducts a
world tour in which the best ranked 34 surfers compete;

{\bf comment:} This only constitutes for male competition

\textcolor{blue}{{\bf response:} We have clarified that we are studying the men's
  championship}

\rule{0mm}{10mm}

{\bf manuscript: } in addition, at each tour venue, two "wildcard" surfers also enter the
competition, who are ignored here

{\bf comment: } Why are they ignored? Especially as some of the wildcards competed in multiple events.

\textcolor{blue}{{\bf response: } Discussed above}

\rule{0mm}{10mm}

{\bf manuscript: } Up to four surfers--the competitors--are in the water simultaneously,
watched by up to five judges.

{\bf comment: } You stated one or two other competitors above. So which is it?

\textcolor{blue}{{\bf response: } Discussed above}

\rule{0mm}{10mm}

{\bf manuscript: } However, for a particular heat, the order statistic--that is, which
competitor scored most highly, second highest, and third--is
informative about competitors abilities: the wave environment is
common to each surfer

{\bf  comment: } But judging is based on the heats conditions, as surfing is performed
within an ever changing environment. Therefore, to try and compare
between surfers throughout a competition is really not applicable. As
each heat, and each wave is different from the previous. So a direct
comparison is somewhat unfair for the competitors.

\textcolor{blue}{{\bf response: } Discussed above}

\rule{0mm}{10mm}

{\bf manuscript: }  and here we apply it to assess the supposed Brazilian preference for
reef-breaking waves.

{\bf comment: } Based on what hypothesis?

\textcolor{blue}{{\bf response: } Discussed above}

### 2022-03-21 inst/surfing_jsa.tex
further work

Surfing is a popular and growing activity in terms of both
participation and viewing \citep{warshaw2010}.  Competitive surfing
involves surfers competing against one or multiple other surfers in heats
lasting 30-50 minutes \citep{booth1995}.  Points are awarded on a
10-point scale: each of five judges awards points to a surfer riding a
particular wave.  The World Surf League (WSL) is the main governing
body for professional surfers \citep{wsl}.  The WSL conducts a world
tour in which the best ranked 34 surfers compete; in addition, at each
tour venue, two ``wildcard" surfers also enter the competition, who
are ignored here due to their generally poor performance and low
participation.

>>>>

Surfing is a popular and growing activity in terms of both
participation and viewing \citep{warshaw2010}.  Competitive surfing
involves surfers competing against one or more other surfers in heats
lasting 30-50 minutes \citep{booth1995}.  Points are awarded on a
10-point scale: each of five judges awards points to a surfer riding a
particular wave.  The World Surf League (WSL) is the main governing
body for professional surfers \citep{wsl}.  The WSL conducts a world
tour in which the best ranked 34 surfers compete; in addition, at each
tour venue, two ``wildcard" surfers also enter the competition, who
are ignored here due to their generally poor performance.
